---
title: "Final Submission Phase1"
author: "Julma Nghiem,Oye Oyegbami,Vishal Raina,Alvina Cheung"
date: "`r Sys.Date()`"
output: html_document
---
##Step 1
##Our group decided to use the raw dataset about salary based on different categorie that we downloaded from kaggle.com. 

##Step 2
##1. The data set our group is going to use is about wages under different categories such as: age, gender, and educational background. 

##2. The data was obtained from the website kaggle.com. https://data.ca.gov/dataset/ca-educational-attainment-personal-income/resource/26201f19-4469-4311-a819-bbbd3e557eda

##3. The team chose this data set because we are interested to see how education level impacts a person's earning ability and if there’s any correlation between the two. As MBA students, it would be interesting how big or small the difference we can expect in terms of salary once we are done with the program. 

library(readr)
library(data.table)
library(readxl)

```{r}
##Step 3
##1. 
group_project_data <- fread ("ca-educational-attainment-personal-income-2008-2014.csv") 
View(group_project_data)
##Is this data tidy? If not, which principles of tidy data does your data set violate?
##2. List the variables and their data types.
group_project_data
class(group_project_data)
mode(group_project_data)
##3. Examine the structure of the data set.
str(group_project_data)
##4. Show the dimensions of the data set. 50 rows and 61 columns. 
nrow(group_project_data)
ncol(group_project_data)
```

```{r}


