---
title: "Final Submission Phase 1"
author: "Julma Nghiem, Oye Oyegbami, Vishal Raina, Alvina Cheung"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(data.table)
library(readxl)
library(dplyr)
library(DataExplorer)
library(tidyr)
library(ggplot2)
```

STEP 1 
   
The data set provides information on orders placed between 2011 and 2014 in 2 countries by 3 segments of consumers:
  (1) Consumer
  (2) Corporate
  (3) Home Office

Each order placed was shipped using one of the 4 shipping modes outlined:
  (1) First Class 
  (2) Same Day
  (3) Second Class
  (4) Standard Class

The order entries originated from 2 countries, United States (US) and Canada (CA). The first two characters of the order ID indicates the country the order was placed. The four digits following the dash separating the country code represents the year the order was placed, and the last 6 digits represents the order ID.  

  STEP 2
  (1)  The data set our group is going to use is about orders under different segments such as: consumer, corporate, and home office.
  (2)  The data was obtained from the website https://foresightbi.com.ng/microsoft-power-bi/dirty-data-samples-to-practice-on/

  (3)  The team chose this data set because we are interested to see how a badly structured dataset can be transformed and cleaned. 
       The orders list has a primary key of Order ID, extracted accros 2 countries, spanning a 5-year period. We would like to analyse the            cleaned data to determine how well each segment is doing as well as track marketing trends by location and shipping modes 

  STEP 3

```{r}

group_project_data <- read_xlsx ("structured-sales-data-1.xlsx") 
View(group_project_data)


plot_missing(group_project_data)

## (1)  The data set is dirty and badly structured. There are some missing value in some column and the Order ID field includes country and            year. 

## (2)  List the variables and their data types.
group_project_data
class(group_project_data)
mode(group_project_data)

## (3)  Examine the structure of the data set.
str(group_project_data)

## (4)  Show the dimensions of the data set. 50 rows and 61 columns. 
nrow(group_project_data)
ncol(group_project_data)

## (5)  Show the first 6 rows and the last 6 rows of data set. 
head(group_project_data)
tail(group_project_data)

## (6)  Identify how many missing values are in your data set?
sum(is.na(group_project_data)) 

## (7)  How many missing values are there in each column?
is.na.data.frame(group_project_data)

## (8) Summarize what cleaning/transforming needs to be done. 
## The rows for the 3 segments need to be aligned and differentiated by their segment names  
## Secondly, the year and country fields need to be extracted from the Order ID.
## Thirdly, the totals and subtotals across each shipping mode computed

```

