---
title: "Final Submission Phase1"
author: "Julma Nghiem,Oye Oyegbami,Vishal Raina,Alvina Cheung"
date: "`r Sys.Date()`"
output: html_document
---
##Step 1
##The group decided to use the raw dataset about salary based on different categorie that we downloaded from kaggle.com. 

##Step 2
##1. The data set our group is going to use is about wages under different categories such as: age, gender, and educational background. 

##2. The data was obtained from the website kaggle.com.  https://data.ca.gov/dataset/ca-educational-attainment-personal-income/resource/26201f19-4469-4311-a819-bbbd3e557eda

##3. The team chose this data set because we are interested to see how education level impacts a person's earning ability and if there’s any correlation between the two. As MBA students, it would be interesting how big or small the difference we can expect in terms of salary once we are done with the program. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(data.table)
library(readxl)
```

##Step 3
##1. 

```{r}
group_project_data <- fread ("ca-educational-attainment-personal-income-2008-2014.csv") 
View(group_project_data)
##The data set is almost tidy. There are some missing value in a column. Also, the format of the column includes date and time. It can be transformed to year only.   
##2. List the variables and their data types.
group_project_data
class(group_project_data)
mode(group_project_data)
##3. Examine the structure of the data set.
str(group_project_data)
##4. Show the dimensions of the data set. 50 rows and 61 columns. 
nrow(group_project_data)
ncol(group_project_data)
##5. Show the first 6 rows and the last 6 rows of data set. 
head(group_project_data)
tail(group_project_data)
##6.Identify how many missing values are in your data set?
sum(is.na(group_project_data)) 
##7. How many missing values are there in each column?
is.na.data.frame(group_project_data)
##8. Summarize what cleaning/transforming needs to be done. 
##First, the Year format needs to be converted to show year only.  Second, the data will be sorted based on the team's interest (for example: by Educational Attainment, by Gender, by Age, and by Personal Income).  Another column can also be created to combine the previously mentioned attributes. 

```

```{r}


