---
title: "Final Submission Phase 1"
author: "Julma Nghiem, Oye Oyegbami, Vishal Raina, Alvina Cheung"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(data.table)
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
```

# -------------
# STEP 1 
# -------------
   
# The data set provides information on orders placed between 2011 and 2014 in 2 countries by 3 segments of # customers:
#  (1) Consumer
#  (2) Corporate
#  (3) Home Office

# Each order placed was shipped using one of the 4 shipping modes outlined:
#   (1) First Class 
#   (2) Same Day
#   (3) Second Class
#   (4) Standard Class

# The order entries originated from 2 countries, United States (US) and Canada (CA). The first two characters of the order ID  indicate the country the order was placed. The four digits following the dash separating the country code represents the year the order was placed, and the last 6 digits represents the order ID.  


# -----------
# STEP 2
# -----------
#  (1)  The data set our group is going to use is about orders under different segments such as: consumer, corporate, and home office.
#  (2)  The data was obtained from the website https://foresightbi.com.ng/microsoft-power-bi/dirty-data-samples-to-practice-on/

#  (3)  The team chose this data set because we are interested to see how a badly structured dataset can be transformed and cleaned. The orders list has a primary key of Order ID, extracted accros 2 countries, spanning a 5-year period. We would like to analyse the cleaned data to determine how well each segment is doing as well as track marketing trends by location and shipping modes.


# -----------
# STEP 3
# -----------
```{r}

group_project_data <- read_xlsx ("structured-sales-data-1.xlsx") 
View(group_project_data)

## (1)  The data set is dirty and badly structured. There are some missing value in some column and the Order ID field includes country and            year. 

## (2)  List the variables and their data types.
group_project_data
class(group_project_data)
mode(group_project_data)

## (3)  Examine the structure of the data set.
str(group_project_data)

## (4)  Show the dimensions of the data set. 50 rows and 61 columns. 
nrow(group_project_data)
ncol(group_project_data)

## (5)  Show the first 6 rows and the last 6 rows of data set. 
head(group_project_data)
tail(group_project_data)

## (6)  Identify how many missing values are in your data set?
sum(is.na(group_project_data)) 

## (7)  How many missing values are there in each column?
is.na.data.frame(group_project_data)

## (8) Summarize what cleaning/transforming needs to be done. 
## The rows for the 3 segments need to be aligned and differentiated by their segment names  
## Secondly, the year and country fields need to be extracted from the Order ID.
## Thirdly, the totals and subtotals across each shipping mode computed

```

  STEP 4 Clean and Tidy the data 

```{r}

### Deletion of columns Consumer Total, Corporate Total, and Home Office Total. 
group_project_data_nototal <- group_project_data %>% select(-`Consumer Total`, -`Corporate Total`, - `Home Office Total`) 
# View(group_project_data_nototal)


### Deletion of row 2 and 3 (Order ID). This row only contains the sub-name or column title for 1st column aka Segment. 
group_project_data_nototal2 <- group_project_data_nototal [-c(1:2,825),]


### Change columns name.  
colnames(group_project_data_nototal2) <- c("Segment>>","Consumer_FirstClass", "Consumer_SameDayClass", "Consumer_SecondClass", "Consumer_StandardClass", "Corporate_FirstClass", "Corporate_SameDayClass", "Corporate_SecondClass", "Corporate_StandardClass", "HomeOffice_FirstClass", "HomeOffice_SameDayClass", "HomeOffice_SecondClass","HomeOffice_StandardClass")
# View(group_project_data_nototal2)


### Gather or PIVOT-LONGER to combine segments: Consumer, Corporate, Home Office. 
group_project_data_nototal3 <- gather(group_project_data_nototal2, key = "Segment", value = "Price", "Consumer_FirstClass":"HomeOffice_StandardClass")
View(group_project_data_nototal3)

### Convert Price (chr) to numberic. 
group_project_data_nototal3$Price <- as.numeric(group_project_data_nototal3$Price)
str(group_project_data_nototal3)


### Remove all NAs in Price column.  
is.na(group_project_data_nototal3)
colSums(is.na(group_project_data_nototal3))
group_project_data_nototal_final <- group_project_data_nototal3[!is.na(group_project_data_nototal3$Price),]


### Separate ORDER ID to country, year, and order ID. 
group_project_data_nototal_final <- separate (group_project_data_nototal_final, "Segment>>", into = c("Country", "Year", "Order ID"), sep = "-")

### For country column, it is either in the United States or Canada.  
group_project_data_nototal_final <- separate (group_project_data_nototal_final, "Segment", into = c("Segment", "Ship Order"), sep = "_")
View(group_project_data_nototal_final)

```

  STEP 5 
  
  (30 points): Perform Exploratory Data Analysis (EDA)
1. Run the summary of the variables in the data set.
2. Learn about the data visually by plotting whichever plot from the list below
applies:
a) Bar plot
b) Histogram
c) Box plot
d) Scatter plot
3. Summarize your findings from the various plots and tables.
4. Uncover new information in the data that is not self-evident (i.e. besides
plotting the data as above, slice and dice the data in different ways, create
new variables, or join separate data frames to create new summary information).

# ------------------------------------------------------
# 1. Run the summary of the variables in the data set.
# ------------------------------------------------------
```{r}


# Total number of orders, total sales by country
Country <-  group_project_data_nototal_final$Country
CountryFreq <- as.data.frame( table(Country) )
 
countryOrderSum <- group_project_data_nototal_final %>% group_by(Country) %>% 
  summarise( sum_country_earned = sum(Price),
            .groups = 'drop')
# countryOrderSum

# orders counts and total sales by Country
CountryOrderPriceTotal <- CountryFreq %>% left_join( countryOrderSum, by='Country' )
CountryOrderPriceTotal

# Canada has more orders than the United states. Approximately 5 times more orders overall and approximately 6 times in sales revenue.
# Need to improve US orders
```

```{r}
ordersByYear <- as.data.frame( table(group_project_data_nototal_final$Year) )
ordersByYear
# Shows the growth trends. Company is doing significantly better with every passing year indicating some level of progress and growth
```

```{r}
ordersBySegment <- as.data.frame( table(group_project_data_nototal_final$Segment) )
ordersBySegment

# Consumer segment is the most successful, followed by Corporate. Efforts should be made to target especially the HomeOffice Segment
```


```{r}
# CANADA

ordersCanada <- subset(completeDatadf, Country=='CA')
ordersCanada

Segment <- ordersCanada$Segment
segmentCountdf <- as.data.frame( table( Segment  ) )

segmentSumCanada <- ordersCanada %>% group_by(Segment) %>% 
  summarise( total_canada_sales = sum(Price),
            .groups = 'drop')

segmentMeanCanada <- ordersCanada %>% group_by(Segment) %>% 
  summarise( mean_canada_sales = mean(Price),
            .groups = 'drop')

allSegmentDataCanada <- segmentCountdf %>% left_join( segmentSumCanada, by='Segment' ) %>% left_join( segmentMeanCanada, by='Segment')

allSegmentDataCanada
```

```{r}
# UNITED STATES

ordersUnitedStates <- subset(completeDatadf, Country=='US')
ordersUnitedStates

Segment <- ordersUnitedStates$Segment
segmentCountdf <- as.data.frame( table( Segment ) )

segmentSumUnitedStates <- ordersUnitedStates %>% group_by(Segment) %>% 
  summarise( total_united_states_sales = sum(Price),
            .groups = 'drop')

segmentMeanUnitedStates <- ordersUnitedStates %>% group_by(Segment) %>% 
  summarise( mean_united_states_sales = mean(Price),
            .groups = 'drop')

allSegmentDataUnitedStates <- segmentCountdf %>% left_join( segmentSumUnitedStates, by='Segment' ) %>% left_join( segmentMeanUnitedStates, by='Segment')

allSegmentDataUnitedStates
```
```{r} 
# UNITED STATES & CANADA

Segment <- group_project_data_nototal_final$Segment
segmentCountdf <- as.data.frame( table( Segment  ) )
completeDatadf <- as.data.frame( group_project_data_nototal_final   )
# 
segmentSum <- completeDatadf %>% group_by(Segment) %>% 
  summarise( sum_dollar_earned = sum(Price),
            .groups = 'drop')
#segmentSum

segmentMean <- completeDatadf %>% group_by(Segment) %>% 
  summarise( mean_dollar_earned = mean(Price),
            .groups = 'drop')
#segmentMean

allSegmentData <- segmentCountdf %>% left_join( segmentSum, by='Segment' ) %>% left_join( segmentMean, by='Segment')

allSegmentData
```

# Unit orders in the HomeOffice Segment are averagely more pricey than other segments.
# Consumer segment had the largest unit order as well as total earned. 



ordersUnitedStates <- subset(completeDatadf, Country=='US')
ordersUnitedStates

```

  STEP 6

  (35 points): Inferences Apply some inferential statistics techniques
covered in Week 10 and in your statistics courses.


```{r}


```
  
  STEP 7

  (35 points):  Predictive Analytics/Modeling Apply some predictive
analytics/modeling techniques covered in Week 10 and in your statistics courses.


```{r}


```
  